# Homework 5
## Anamika Das
========================================================


```{r, echo = FALSE, include=FALSE}
packages = c("MASS","ggplot2","arm", "AER","dplyr","foreign")             
lapply(packages, library, character.only = TRUE)

```


```{r, echo = FALSE, include=FALSE}
setwd("~/Documents/UW")
```


The backward and forward stepwise regression produce the same model with same variables and coefficients.


```{r, results='hide'}

abal<-read.csv("abaloneTrain.csv")
abalmod<-lm(rings~(length+width+height+meatweight+shellweight)^2,data=abal)
mod.back = stepAIC(object = abalmod, direction = "backward")

abalnull = lm(rings ~ 1, data = abal)
mod.for = stepAIC(object = abalnull, scope = ~ (length+width+height+meatweight+shellweight)^2,
                  direction = "forward")

```

```{r}
mod.for
```

Removing the outlier lowered the AIC of the model. 
```{r}
#to remove outlier
abal[abal$height>1, ]
abal2<-abal[-1217, ]
```

```{r, results='hide'}
abalmod2<-lm(rings~(length+width+height+meatweight+shellweight)^2,data=abal2)
mod.back2 = stepAIC(object = abalmod2, direction = "backward")


abalnull2 = lm(rings ~ 1, data = abal2)
mod.for2 = stepAIC(object = abalnull2, scope = ~ (length+width+height+meatweight+shellweight)^2,
                  direction = "forward")
```

With outlier removed
```{r}
mod.for2
```


G&H Question1
```{r}

shotsdist<-function(n){
  a<-1
  trial<-vector() 
  for(a in 1:n){  
      score<-vector()
      i<-1     
      score[i]<-rbinom(1,1,.6)
       i<-i+1
      score[i]<-rbinom(1,1,.6)
        while(sum(score[i],score[i-1])>0){
          i<-i+1
          score[i]<-rbinom(1,1,.6)            
        }  
    trial[a]<-length(score)
    a<-a+1
    }
  return(trial)
}

simshots<-shotsdist(100000)
c(mean=mean(simshots),sd=sd(shotsdist(1000)))     
hist(shotsdist(1000),freq=FALSE)

```

c. 
```{r}

shotstable<-function(n){
      table<-data.frame()  
      a<-1
      trial<-vector() 
      for(a in 1:n){  
          score<-vector()
          i<-1     
          score[i]<-rbinom(1,1,.6)
           i<-i+1
          score[i]<-rbinom(1,1,.6)
            while(sum(score[i],score[i-1])>0){
              i<-i+1
              score[i]<-rbinom(1,1,.6)            
            }  
        table[a,1]<-length(score)
        table[a,2]<-sum(score)
        names(table)<-c("attempts", "successes")
        a<-a+1
        }
      return(table)
}

d1<-shotstable(1000)
plot(d1$attempts,d1$successes)
plot(d1$attempts,d1$successes/d1$attempts)

```


(d)

We fail to reject the null hypothesis. The new player is not better than the original player.
```{r}
(15-mean(d1$attempts))/sd(d1$attempts)
```

G&H Question2

```{r}
    #n1=number of males
    #n2=number of females
i<-1
pbreak<-vector()
for (i in 1:1000){
    n1<-rbinom(1,10,.48)
    n2<-10-n1
    mw<-rnorm(n1,exp(5.13),exp(.17))
    fw<-rnorm(n2,exp(4.96),exp(.2))
    ifelse(sum(mw,fw)>1650,pbreak[i]<-1,pbreak[i]<-0)
    i<-i+1
}

sum(pbreak)/1000

````

G&H Question4
```{r}
data(TeachingRatings)
mod<-lm(eval~gender+beauty+age+native, data=TeachingRatings) 
mod1<-lm(eval~gender+beauty+age+native+native:age, data=TeachingRatings)
display(mod)



ceval<-coef(mod)
preda<- ceval[1]+ceval[2]*1+ceval[3]*0+ceval[4]*0
predb<-ceval[1]+ceval[3]*-.5+ceval[4]*0
names(preda)<-NULL
names(predb)<-NULL
#Resdiual sd=.53
inta<-rnorm(1000,preda,.53)
intb<-rnorm(1000,predb,.53)
pred.diff<-inta-intb
hist(pred.diff,freq=FALSE)
```

Probability Instuctor A has a higher rating

```{r}
diff<-vector()
i<-1
for (i in 1:length(pred.diff))  {
  ifelse(pred.diff[i]>0, diff[i]<-1,diff[i]<-0)
  i<-i+1
}
sum(diff)/length(diff)

```